A Dependently Typed Multi-Stage Calculus\thanks{Supported by organization x.}

### 2019-05-28 main.tex
Fix by grammarly.

% 多段階計算のルール
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp are rules for a multistage calculus.
They are corresponding to quoting a code, unquoting a code, making a stage abstraction, 
application of $\epsilon$ stage to a stage abstraction, and cross-stage persistence, respectively.
Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

% \TConv
The main technical point of this paper is \TConv\ and type equivalence rules needed in \TConv.
This rule allows ut to replace a type with another type that is equivalent.
In a type system which includes dependent types, this kind of rule is essential
because two types which have different shapes may be equivalent.
For example, when we use a matrix type which have their sizes ($\textrm{Mat}\ n\ m$),
$\textrm{Mat}\ 5\ 3$ is equivalent to $\textrm{Mat}\ (4+1)\ (1+2)$ obviously.

@@

% \QCsp以外の説明
We show type equivalence rules in Figure \ref{fig:type-equivalence-rules}.
All rules except \QTRefl, \QTSym, \QTTrans, and \QTApp\ are generated naturally from the typing rules.
\QTRefl, \QTSym, \QTTrans\ exist in order to make the type equivalence relation an equivalence relationship.
The rule \QTApp means that if there are two equivalent $\Pi$ type and two equivalent terms,
the result of applications are also equivalent.
A judgement of $\G \V M \E N : \rho @ A$ means that $M$ and $N$ are equivalent.

@@

% \QPercentの説明
The rule \QPercent is something special because it has no background of syntax or reductions.
But we add this rule to \LMD\ because it is essential when you use dependent types in code type.
For example, we can utilize a dependently typed multistage calculus to generate a code for matrix multiplication.
When we know sizes of matrices before the multiplication,
we can generate efficient code for multiplication by unrolling loops.

$\text{mulmat}$ function in following code fragment generate a code for matrix multiplication.
$\text{mulmat}$ takes two integer which are the size of multiplicand matrix and generate a code.
The last integer to decide the size of multiplier matrix is given at runtime.
You can generate a code by applying two integers to $\text{mulmat}$.
We applied 3 and 5 in the second line and got $\TW_\alpha \Pi z:\I.(\M\ z\ \%_\alpha 5) \to (\M\ \%_\alpha 5\ \%_\alpha 3) \to (\M\ z\ \%_\alpha 3)$.
But this type is difficult to combine with other code because there are two $\%_\alpha$ for CSP.
\QPercent\ states that we can erase this $\%_\alpha$ under a conditions which is explained in the next paragraph.

@@

This kind of type equality is very useful when we combine $\text{mulmat}\ 3\ 5$ with another code.
We cannot remove $\%_\alpha$ in the type without \QPercent and it makes very difficult to combine
because types of codes don't contains $\%_\alpha$ generally.

The reduction given above is a full reduction and any redexes can be reduced in arbitrary order.
But we need to fix a order of reduction when we implement a interpreter of \LMD and
the order must be appropriate in terms of stages.

@@

Values at $\epsilon$ stage are a $\lambda$ abstraction, a quoted code, or a $\Lambda$ abstraction.
The body of a $\lambda$ abstraction can be any term but the body of $\Lambda$ abstraction only can be a value.
This restriction means that the body of $Lambda$ abstraction must evaluated before application of an $\epsilon$ stage.
Values at non-$\epsilon$ stage includes all term except $\TBL_\alpha v^\epsilon$.
This is because $\TBL_\alpha v^\epsilon$ is a redex.

@@

Substitution lemma for \LMD\ is little more complicated than ordinary one 
because there are six types of judgement and two types of substitution in \LMD.

>>>>

% 多段階計算のルール
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp\ are rules for a multistage calculus.
They are corresponding to quoting a code, unquoting a code, making a stage abstraction, 
application of $\epsilon$ stage to a stage abstraction, and cross-stage persistence, respectively.
Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

% \TConv
The main technical point of this paper is \TConv\ and type equivalence rules needed by \TConv.
This rule allows ut to replace a type with another type that is equivalent.
In a type system which includes dependent types, this kind of rule is essential
because two types which have different shapes may be equivalent.
For example, when we use a matrix type which have their sizes ($\textrm{Mat}\ n\ m$),
$\textrm{Mat}\ 5\ 3$ is equivalent to $\textrm{Mat}\ (4+1)\ (1+2)$ obviously.

@@

% \QCsp以外の説明
We show type equivalence rules in Figure \ref{fig:type-equivalence-rules}.
All rules except \QTRefl, \QTSym, \QTTrans, and \QTApp\ are generated naturally from the typing rules.
\QTRefl, \QTSym, \QTTrans\ exist in order to make the type equivalence relation an equivalence relationship.
The rule \QTApp\ means that if there are two equivalent $\Pi$ type and two equivalent terms,
the results of applications are also equivalent.
A judgement of $\G \V M \E N : \rho @ A$ means that $M$ and $N$ are equivalent.

@@

% \QPercentの説明
The rule \QPercent is something special because it has no background of syntax or reductions.
But we add this rule to \LMD\ because it is essential when you use dependent types in code type.
For example, we can utilize a dependently typed multistage calculus to generate code for matrix multiplication.
When we know sizes of matrices before the multiplication,
we can generate efficient code for multiplication by unrolling loops.

$\text{mulmat}$ function in the following code fragment generate code for matrix multiplication.
$\text{mulmat}$ takes two integers which are the size of the multiplicand matrix and generate a code.
The last integer to decide the size of the multiplier matrix is given at runtime.
You can generate code by applying two integers to $\text{mulmat}$.
We applied 3 and 5 in the second line and got $\TW_\alpha \Pi z:\I.(\M\ z\ \%_\alpha 5) \to (\M\ \%_\alpha 5\ \%_\alpha 3) \to (\M\ z\ \%_\alpha 3)$.
But this type is difficult to combine with other code because there are two $\%_\alpha$ for CSP.
\QPercent\ states that we can erase this $\%_\alpha$ under a condition which is explained in the next paragraph.

@@

This kind of type equality is very useful when we combine $\text{mulmat}\ 3\ 5$ with another code.
We cannot remove $\%_\alpha$ in the type without \QPercent and it makes very difficult to combine
because types of codes don't contain $\%_\alpha$ generally.

The reduction given above is a full reduction and any redexes can be reduced in arbitrary order.
But we need to fix order of reduction when we implement an interpreter of \LMD and
the order must be appropriate in terms of stages.

@@

Values at $\epsilon$ stage are a $\lambda$ abstraction, a quoted code, or a $\Lambda$ abstraction.
The body of a $\lambda$ abstraction can be any term but the body of $\Lambda$ abstraction only can be a value.
This restriction means that the body of $Lambda$ abstraction must be evaluated before application of an $\epsilon$ stage.
Values at non-$\epsilon$ stage includes all term except $\TBL_\alpha v^\epsilon$.
This is because $\TBL_\alpha v^\epsilon$ is a redex.

@@

Substitution lemma for \LMD\ is little more complicated than an ordinary one 
because there are six types of judgment and two types of substitution in \LMD.
